Understanding the impacts of video-guided activities on parent-child interaction

### 2020-04-01 paper/manuscript.tex
final stats check

\shorttitle{Impacts of videos on parent-child interaction}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Stanford University\\\textsuperscript{2} Gallaudet University\\\textsuperscript{3} Cornell University}
\keywords{digital parenting advice; joint attention; lexical diversity; guided play\newline\indent Word count: 4629}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

We fit a Bayesian mixed-effects linear regression predicting TTR as a function of condition, age (centered), and their interaction with a random intercept per video using rstanarm (Goodrich, Gabry, Ali, \& Brilleman, 2018).
For effects that are at least 95\% likely to be non-zero according to the posterior distribution (i.e., probability of direction: \emph{pd}; Makowski, Ben-Shachar, Chen, \& Lüdecke (2019)), we report estimated coefficients (\(\beta\)) as well as 89\% Bayesian credible intervals (89\% CI)\footnote{89\% CIs are recommended for Bayesian analyses because unless the effective sample size (ESS) is on the order of 10,000, the 95\% credible interval is unstable (Kruschke, 2014; McElreath, 2018). Our ESS is approximately 2,100, and thus we go with the tighter, more stable interval.}, demarcating the range within which 89\% of the posterior falls, meaning that given the observed data, the effect has 89\% probability of falling within this range.
There was lower TTR in the Activity Video condition (mean: 0.10) than in the No Video condition (mean: 0.14, \(\beta=\)-0.04, 89\% CI={[}-0.052, -0.027{]}, \emph{pd}= 1).
A similar regression instead predicting MTLD also found lower lexical diversity in the Activity Video condition (mean MTLD: 5.46) than in the No Video condition (mean: 8.49, \(\beta=\)-3.04, 89\% CI={[}-4.505, -1.448{]}, \emph{pd}= 0.998{]}), with no predicted influence of age.
Figure \ref{fig:lexdiv} shows the mean of each lexical diversity measure (TTR and MTLD) by condition.

We also conducted similar regressions predicting the rate of word tokens and types per minute of play, finding an effect of condition on the rate of word tokens (\(\beta=\) 16.50, 89\% CI={[}5.778, 27.201{]}, \emph{pd}= 0.994), with parents using tokens at a higher rate in the Activity Video condition (mean: 69 tokens/min, bootstrapped 95\% confidence interval (conf. int.): {[}193, 259{]}) than in the No Video condition (mean: 52, 95\% conf. int.: {[}139, 194{]}).

@@

We fit a Bayesian mixed-effects linear regression predicting the rate of parent bids (per minute) for joint attention (JA) as a function of fixed effects of condition, age (centered), and their interaction, with random intercepts per video.
Parents' bid rate was greater in the Activity Video condition (mean: 1.83, sd: 0.79) than in the No Video condition (mean: 1.10, sd: 0.83, \(\beta=0.71\), 89\% CI={[}0.27, 1.16{]}, \emph{pd}= 0.99).
Mixed-effects regressions with the same structure were performed predicting the rate of episodes of coordinated and passive JA, and the time spent in coordinated and passive JA (seconds per minute of play).
There were no notable effects on the rate of episodes nor on the time spent in coordinated or passive JA episodes.
Figure \ref{fig:JA} (top) shows the mean rate of bids and episodes of JA by condition in Experiment 1.

@@

We also fit Bayesian mixed-effects linear regression models predicting each of the above lexical diversity and joint attention dependent variables as a function of fixed effects of condition, age (centered), gender, parent's education level, and the subscales of the EPAQ: Early Learning (EL), Affection and Attachment (AA), and Rules and Respect (RR), along with interactions of condition and EL, AA, and RR.
These models included random intercepts per video.
Of these exploratory regressions, only the regression predicting the rate of passive JA episodes showed an effect involving these demographic variables: parents scoring higher on the Rules and Respect (RR) subscale had a lower rate of passive JA episodes (\(\beta=NA\), 89\% CI={[}-0.599, -0.078{]}, \emph{pd}= 0.98).
However, higher RR parents in the Activity Video condition did not show a decrease in passive JA (interaction of condition and RR: \(\beta=0.60\), 89\% CI={[}0.221, 0.955{]}, \emph{pd}= 0.992).

To better understand the intervention's effect on language use, we analyzed which words were characteristic of parents' speech in each condition, comparing the difference in frequency rank of each word (lemma) in the two conditions, as well as contrasting the corpus overall with a general English-language word frequency list (see SI for the interactive corpus characteristic plot).
Words that were strongly indicative of being from the Activity Video condition include \enquote{give}, \enquote{big}, \enquote{small}, \enquote{ribbit}, \enquote{thank}, \enquote{have}, and \enquote{bus}, while words that were most characteristic of the No Video condition include \enquote{shake}, \enquote{ready}, \enquote{oh}, \enquote{on}, \enquote{going}, \enquote{going}, \enquote{see}, \enquote{let}, and the child's name.

@@

In summary, while parents produced more word types and tokens after viewing the activity video, lexical diversity (both TTR and MTLD) was higher when parents were just asked to play as they normally would.
It may be that parents in the Activity Video condition, in their attempt to stick to the prescribed task, end up repeating themselves more, and indeed some differences in speech acts were notable:
after the Activity Video, parents used more words related to requests (e.g., \enquote{Can I have X? / Give me X. Thank you!}), whereas after no intervention parents' language related more to invitations (e.g., \enquote{Are you ready?} / \enquote{Let's see.}).
However, parents who watched an activity video also made more bids for JA with their child.
While this did not result in a greater number of successful episodes of JA--passive or coordinated--than dyads in the No Video condition, it should be noted that low reliability for coding of passive JA episodes may have limited our power to find an effect, leading us to refine our coding procedure for Experiment 2.
In sum, the results of Experiment 1 suggest that digital parenting advice can increase parents' efforts to engage their child in joint attention, expand the volume if not diversity of their speech, and can shift the type of speech acts towards more requests.

@@

Experiment 1 found that parents who watched an activity video made more bids for joint attention and spoke more words overall to their children, but had lower lexical diversity compared to parents who played with their children as they normally would at home.
Might it be that parents who are focused on a specific activity show reduced lexical diversity due to their focus on engaging their child in the activity?
Experiment 2 focuses on replicating the key findings using a stronger control group, as well as a restricted number of preregistered predictions.\footnote{Preregistration: \href{https://osf.io/2bpdf}{https://osf.io/2bpdf/}.}

@@

Parents' child-directed speech during the play sessions (mean duration: 3.08 min) was transcribed and processed, and bids and episodes of joint attention were coded according to the same procedure used in Experiment 1.
We first report preregistered regressions\footnote{Although the preregistration implied the use of standard linear mixed-effects regression through the specification of adopting an alpha level of .005 for statistical significance, the non-convergence of some of the below regressions led us to switch to Bayesian regression. Using a Bayesian analysis also has the added benefit of not requiring arbitrary decisions about changing alpha levels to attempt to correct for multiple comparisons (Gelman, 2008).} predicting TTR and number of tokens per minute of play, as well as an exploratory regression predicting MTLD per minute.
We then turn to preregistered regressions of parental bids for joint attention and the total number of JA episodes.

@@

We fit a Bayesian mixed-effects linear regression predicting TTR as a function of age (centered) and condition with an interaction term, and with random intercepts per video.
This revealed lower TTR after the Activity Video (mean: 0.12) than after the Science Video (mean: 0.15, \(\beta=-0.03\), , 89\% CI={[}-0.046, -0.014{]}, \emph{pd}= 0.997).
The preregistered regression predicting the number of tokens used by parents revealed no effects.
An exploratory mixed-effects linear regression predicting MTLD found no effect of age or condition.
Figure \ref{fig:lexdiv} (bottom left and middle) shows the mean of each lexical diversity measure (TTR and MTLD) by condition.
Regressions with the same structure predicting the number of words tokens found no effect of age or condition.
The means of the lexical measures are shown in Table \ref{e2tab}.

@@

\caption{\label{tab:e2tab}\label{e2tab} Lexical diversity measures in Experiment 2.}
\centering
\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
\hline
Condition & TTR (M) & (sd) & MTLD (M) & (sd) & Types (M) & (sd) & Tokens (M) & (sd)\\
\hline
Science Video & 0.15 & 0.03 & 7.28 & 3.41 & 28.05 & 8.54 & 61.85 & 23.58\\
\hline
Activity Video & 0.12 & 0.04 & 6.74 & 2.80 & 26.57 & 8.39 & 77.22 & 32.09\\
\hline
\end{tabular}
\end{table}

@@

We fit Bayesian mixed-effects linear regressions predicting the rate of parental bids for joint attention and the rate of JA episodes as a function of fixed effects of condition, age (centered), and their interaction, with random intercepts per video.
Shown in Figure \ref{fig:JA} (left bottom), parents made bids for JA at a greater rate after watching the Activity Video (mean: 4.20, 95\% conf. int.: {[}3.81, 4.63{]}; \(\beta=1.02\), 89\% CI={[}0.51, 1.53{]}, \emph{pd}= 1)) than after the Science Video (mean: 3.18, 95\% conf. int.: {[}2.81, 3.54{]}).
There were no other effects on parental bids for JA.

The regression predicting rate of JA episodes revealed an effect of condition (\(\beta=0.50\), e2tjaCond, with JA episodes occurring at a greater rate after the Activity Video (mean: 2.86, 95\% conf. int.: {[}2.61, 3.15{]}) than after the Control Video (mean: 2.36, 95\% conf. int.: {[}2.13, 2.61{]}).
Older children also participated in JA episodes at a lesser rate than younger children (\(\beta=-1.55\), 89\% CI={[}-2.46, -0.63{]}, \emph{pd}= 0.99).
However, this age effect was moderated in the Activity Video condition (\(\beta=1.70\), 89\% CI={[}0.33, 2.99{]}, \emph{pd}= 0.97): shown in Figure \ref{fig:JA} (right bottom), older children did not engage in episodes of JA at a lower rate after an activity video.

@@

Four additional exploratory regressions with a similar structure were carried out to predict the number and duration of coordinated and passive JA episodes.
The regression predicting the rate of coordinated JA episodes found an effect of condition (\(\beta=0.46\), 89\% CI={[}0.143, 0.769{]}, \emph{pd}= 0.989), with a greater rate of coordinated JA episodes occurring after the Activity Video (mean: 2.18, 95\% CI: {[}1.89, 2.47{]}) than after the Control Video (mean: 1.72, 95\% CI: {[}1.53, 1.94{]}).
There was an interaction of age and condition (\(\beta=1.60\), 89\% CI={[}0.405, 2.874{]}, \emph{pd}= 0.984), shown in Figure \ref{fig:e2ja-coord}, revealing that after an Activity Video older children participated in coordinated JA episodes at a greater rate than children in the Control Video condition.
The regression predicting the time spent in coordinated JA episodes found no notable effects.

Older children both engaged in episodes of passive JA at a greater rate with their caregiver (\(\beta=-0.88\), 89\% CI={[}-1.514, -0.166{]}, \emph{pd}= 0.972), and spent more time in passive JA with their caregiver (\(\beta=-0.01\), 89\% CI={[}-0.01, -0.002{]}, \emph{pd}= 0.99).
Overall, these results show that the older children in our sample engage in more and longer episodes of joint attention with their caregivers, and that activity videos in particular lead to more episodes of coordinated JA.

@@

\leavevmode\hypertarget{ref-Makowski2019}{}%
Makowski, D., Ben-Shachar, M. S., Chen, S. H. A., \& Lüdecke, D. (2019). Indices of effect existence and significance in the bayesian framework. \emph{Frontiers in Psychology}, \emph{10}, 2767. \url{http://doi.org/10.3389/fpsyg.2019.02767}

>>>>

\shorttitle{Impacts of videos on parent-child interaction}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Stanford University\\\textsuperscript{2} Gallaudet University\\\textsuperscript{3} Cornell University}
\keywords{digital parenting advice; joint attention; lexical diversity; guided play\newline\indent Word count: 5056}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

We fit a Bayesian mixed-effects linear regression predicting TTR as a function of condition, age (centered), and their interaction with a random intercept per video using rstanarm (Goodrich, Gabry, Ali, \& Brilleman, 2018).
For effects that are at least 95\% likely to be non-zero according to the posterior distribution (i.e., probability of direction: \emph{pd}; Makowski et al. (2019a)), we report estimated coefficients (\(\beta\)) as well as 89\% Bayesian credible intervals (89\% CI)\footnote{89\% CIs are recommended for Bayesian analyses because unless the effective sample size (ESS) is on the order of 10,000, the 95\% credible interval is unstable (Kruschke, 2014; Makowski et al., 2019b; McElreath, 2018). Our ESS ranges from 2,000-5,000, and thus we report the tighter, more stable interval.}, demarcating the range within which 89\% of the posterior falls, meaning that given the observed data, the effect has 89\% probability of falling within this range (see Appendix for more an illustrated explanation).
There was lower TTR in the Activity Video condition (mean: 0.32) than in the No Video condition (mean: 0.43, \(\beta=\)-0.11, 89\% CI = {[}-0.15, -0.07{]}, \emph{pd} = 1).
A similar regression instead predicting MTLD also found lower lexical diversity in the Activity Video condition (mean MTLD: 17.87) than in the No Video condition (mean: 27.09, \(\beta=\)-9.24, 89\% CI = {[}-14.07, -4.21{]}, \emph{pd} = 1{]}), with no notable influence of age.
Figure \ref{fig:lexdiv} (top) shows the mean of each lexical diversity measure (TTR and MTLD) by condition.

We also conducted similar regressions predicting the rate of word tokens and types per minute of play, finding a notable effect of condition on the rate of word tokens (\(\beta=\) 16.50, 89\% CI = {[}5.78, 27.2{]}, \emph{pd} = 0.99), with parents using tokens at a higher rate in the Activity Video condition (mean: 69 tokens/min, bootstrapped 95\% confidence interval (conf. int.): {[}60, 79{]}) than in the No Video condition (mean: 52, 95\% conf. int.: {[}44, 61{]}).

@@

We fit a Bayesian mixed-effects linear regression predicting the rate of parent bids (per minute) for joint attention (JA) as a function of fixed effects of condition, age (centered), and their interaction, with random intercepts per video.
Parents' bid rate was greater in the Activity Video condition (mean: 3, sd: 1.11) than in the No Video condition (mean: 2.15, sd: 1.02, \(\beta=0.72\), 89\% CI = {[}0.12, 1.32{]}, \emph{pd} = 0.97).
Mixed-effects regressions with the same structure were performed predicting the rate of episodes of coordinated and passive JA, and the time spent in coordinated and passive JA.
There were no notable effects on the rate of episodes nor on the time spent in coordinated or passive JA episodes.
Figure \ref{fig:JA} (top) shows the mean rate of bids and episodes of JA by condition in Experiment 1.

@@

We also fit Bayesian mixed-effects linear regression models predicting each of the above lexical diversity and joint attention dependent variables as a function of fixed effects of condition, age (centered), the child's sex, parent's education level, and the subscales of the EPAQ: Early Learning (EL), Affection and Attachment (AA), and Rules and Respect (RR), along with interactions of condition and EL, AA, and RR.
These models included random intercepts per video.
Of these exploratory regressions, two showed notable effects involving the EPAQ subscales, and one other showed an effect of the child's sex.
First, in the word types regression, parents with higher Affection and Attachment scores used more word types per minute after watching an Activity Video (interaction of condition and AA: \(\beta=6.81\), 89\% CI = {[}0.87, 13.13{]}, \emph{pd} = 0.97).
Second, in the regression examining the rate of passive JA episodes, parents scoring higher on the Rules and Respect (RR) subscale had a lower rate of passive JA episodes (\(\beta=-0.34\), 89\% CI = {[}-0.62, -0.07{]}, \emph{pd} = 0.98).
However, higher RR parents who saw an Activity Video did not show this decrease in passive JA (interaction of condition and RR: \(\beta=0.63\), 89\% CI = {[}0.23, 1.03{]}, \emph{pd} = 0.99).
This regression also found that more educated parents showed a lower rate of passive JA episodes (\(\beta=-0.25\), 89\% CI = {[}-0.45, -0.04{]}, \emph{pd} = 0.97).
Finally, parents made bids for JA at a higher rate with male children than with female children (male mean: 2.76 bids/min, female mean: 2.49, \(\beta=0.82\) 89\% CI = {[}0.06, 1.54{]}, \emph{pd} = 0.96).

To get a better sense of the intervention's effect on language use, we analyzed which words were characteristic of parents' speech in each condition, comparing the difference in frequency rank of each word (lemma) in the two conditions, as well as contrasting the corpus overall with a general English-language word frequency list (see SI for the interactive corpus characteristic plot).
Words that were strongly indicative of being from the Activity Video condition include \enquote{give}, \enquote{big}, \enquote{small}, \enquote{ribbit}, \enquote{thank}, \enquote{have}, and \enquote{bus}, while words that were most characteristic of the No Video condition include \enquote{shake}, \enquote{ready}, \enquote{oh}, \enquote{on}, \enquote{going}, \enquote{going}, \enquote{see}, \enquote{let}, and the child's name.

@@

In summary, while parents produced more word types and tokens after viewing the activity video, lexical diversity (both TTR and MTLD) was higher when parents were just asked to play as they normally would.
It may be that parents in the Activity Video condition, in their attempt to stick to the prescribed task, end up repeating themselves more, and indeed some differences in speech acts were notable:
after the Activity Video, parents used more words related to requests (e.g., \enquote{Can I have X? / Give me X. Thank you!}), whereas after no intervention parents' language related more to invitations (e.g., \enquote{Are you ready?} / \enquote{Let's see.}).
However, parents who watched an activity video also made more bids for JA with their child.
This did not result in a greater number of successful episodes of JA---passive or coordinated---than dyads in the No Video condition, although low reliability in passive JA coding (which led us to refine our JA coding guidelines for Experiment 2) may limit our ability to detect an effect there.
In sum, the results of Experiment 1 suggest that digital parenting advice can increase parents' efforts to engage their child in joint attention, expand the volume if not diversity of their speech, and can shift the type of speech acts towards more requests.

@@

Experiment 1 found that parents who watched an activity video made more bids for joint attention and spoke more words tokens per minute to their children, but had lower lexical diversity compared to parents who played with their children as they normally would at home.
Might it be that parents who are focused on a specific activity show reduced lexical diversity due to their focus on engaging their child in the activity?
Experiment 2 focuses on replicating the key findings using a stronger control group, as well as a restricted number of preregistered predictions.\footnote{Preregistration: \href{https://osf.io/2bpdf}{https://osf.io/2bpdf/}.}

@@

Parents' child-directed speech during the play sessions (mean duration: 3.08 min) was transcribed and processed, and bids and episodes of joint attention were coded according to the same procedure used in Experiment 1.
We first report preregistered regressions\footnote{Although the preregistration implied the use of standard linear mixed-effects regression through the specification of adopting an alpha level of .005 for statistical significance, the non-convergence of some of the below regressions led us to switch to Bayesian regression. Using a Bayesian analysis also has the added benefit of not requiring arbitrary decisions about changing alpha levels to attempt to correct for multiple comparisons (Gelman, 2008).} predicting TTR and rate of word tokens, as well as an exploratory regression predicting MTLD.
We then turn to preregistered regressions of parental bids for joint attention and the total number of JA episodes.

@@

We fit a Bayesian mixed-effects linear regression predicting TTR as a function of age (centered) and condition with an interaction term, and with random intercepts per video.
This revealed lower TTR after the Activity Video (mean: 0.38) than after the Science Video (mean: 0.48, \(\beta=-0.09\), , 89\% CI = {[}-0.14, -0.05{]}, \emph{pd} = 1).
The preregistered regression predicting the number of tokens used by parents revealed no effects.
An exploratory mixed-effects linear regression predicting MTLD found no effect of age or condition.
Figure \ref{fig:lexdiv} (bottom left and middle) shows the mean of each lexical diversity measure (TTR and MTLD) by condition.
Regressions with the same structure predicting the number of words tokens found no effect of age or condition.
The means of the lexical measures are shown in Table \ref{e2tab}.

@@

\caption{\label{tab:e2tab}\label{e2tab} Lexical diversity measures in Experiment 2.}
\centering
\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
\hline
Condition & TTR (M) & (sd) & MTLD (M) & (sd) & Types (M) & (sd) & Tokens (M) & (sd)\\
\hline
Science Video & 0.48 & 0.08 & 22.45 & 10.57 & 28.05 & 8.54 & 61.85 & 23.58\\
\hline
Activity Video & 0.38 & 0.12 & 20.63 & 8.66 & 26.57 & 8.39 & 77.22 & 32.09\\
\hline
\end{tabular}
\end{table}

@@

We fit Bayesian mixed-effects linear regressions predicting the rate of parental bids for joint attention and the rate of JA episodes as a function of fixed effects of condition, age (centered), and their interaction, with random intercepts per video.
Shown in Figure \ref{fig:JA} (left bottom), parents made bids for JA at a greater rate after watching the Activity Video (mean: 4.23, 95\% conf. int.: {[}3.85, 4.65{]}; \(\beta=1.05\), 89\% CI = {[}0.53, 1.58{]}, \emph{pd} = 1)) than after the Science Video (mean: 3.19, 95\% conf. int.: {[}2.83, 3.57{]}).
There were no other effects on parental bids for JA.

The regression predicting rate of JA episodes revealed an effect of condition (\(\beta=0.51\), e2tjaCond, with JA episodes occurring at a greater rate after the Activity Video (mean: 2.88, 95\% conf. int.: {[}2.61, 3.17{]}) than after the Control Video (mean: 2.36, 95\% conf. int.: {[}2.12, 2.60{]}).
Older children also participated in JA episodes at a lesser rate than younger children (\(\beta=-1.57\), 89\% CI = {[}-2.42, -0.65{]}, \emph{pd} = 1).
However, this age effect was moderated in the Activity Video condition (\(\beta=1.69\), 89\% CI = {[}0.34, 2.98{]}, \emph{pd} = 0.98): shown in Figure \ref{fig:JA} (right bottom), older children did not engage in episodes of JA at a lower rate after an activity video.

@@

Four additional exploratory regressions with a similar structure were carried out to predict the number and duration of coordinated and passive JA episodes.
The regression predicting the rate of coordinated JA episodes found an effect of condition (\(\beta=0.46\), 89\% CI = {[}0.13, 0.77{]}, \emph{pd} = 0.99), with a greater rate of coordinated JA episodes occurring after the Activity Video (mean: 2.19, 95\% CI: {[}1.93, 2.46{]}) than after the Control Video (mean: 1.72, 95\% CI: {[}1.52, 1.93{]}).
There was an interaction of age and condition (\(\beta=1.59\), 89\% CI = {[}0.35, 2.94{]}, \emph{pd} = 0.98), shown in Figure \ref{fig:e2ja-coord}, revealing that after an Activity Video older children participated in coordinated JA episodes at a greater rate than children in the Control Video condition.
The regression predicting the time spent in coordinated JA episodes found no notable effects.

Older children both engaged in episodes of passive JA at a greater rate with their caregiver (\(\beta=-0.89\), 89\% CI = {[}-1.54, -0.2{]}, \emph{pd} = 0.98), and spent more time in passive JA with their caregiver (\(\beta=-5.92\), 89\% CI = {[}-9.85, -1.83{]}, \emph{pd} = 0.99).
Overall, these results show that the older children in our sample engage in more and longer episodes of joint attention with their caregivers, and that activity videos in particular lead to more episodes of coordinated JA.

@@

\leavevmode\hypertarget{ref-Makowski2019}{}%
Makowski, D., Ben-Shachar, M. S., Chen, S. H. A., \& Lüdecke, D. (2019a). Indices of effect existence and significance in the bayesian framework. \emph{Frontiers in Psychology}, \emph{10}, 2767. \url{http://doi.org/10.3389/fpsyg.2019.02767}

\leavevmode\hypertarget{ref-bayestestR}{}%
Makowski, D., Ben-Shachar, M. S., \& Lüdecke, D. (2019b). BayestestR: Describing effects and their uncertainty, existence and significance within the bayesian framework. \emph{Journal of Open Source Software}, \emph{4}(40), 1541. \url{http://doi.org/10.21105/joss.01541}
